---
title: "Whats going on in the Delta Wateshed"
---

------------------------------------------------------------------------

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
rm(list=ls())


library(rvest)

library(pdftools)

library(tidyverse)

library(stringr)

library(rvest)

library(xml2)

library(CDECRetrieve)

library(magrittr)

library(leaflet)

library(writexl)

library(readxl)

library(lubridate)

library(openxlsx)

library(geojsonio)

library(htmltools)

```

```{r}

# Read PDF text directly from the URL without saving
northern_pdf_url <- "https://cdec.water.ca.gov/cgi-progs/products/PLOT_ESI.pdf"
northern_pdf_text <- pdf_text(northern_pdf_url)

#can print if want to debugg pdf
#print(northern_pdf_text)

```

# Precipitation

#### Northern Sierra Region Current precipitation {style="color: darkblue"}

```{r}

# Read PDF text directly from the URL
northern_pdf_url <- "https://cdec.water.ca.gov/cgi-progs/products/PLOT_ESI.pdf"
northern_pdf_text <- pdf_text(northern_pdf_url)

# Combine all pages into a single string (ensures we search the entire document at once)
combined_text <- paste(northern_pdf_text, collapse = " ")

# Define regex pattern to extract the number next to "Current:
#"regex (regular expression) pattern is a sequence of characters that define a search pattern for text. to extract the number next to "Current:"

pattern <- "(?i)Current:\\s?(\\d{1,4}(\\.\\d{1,2})?)"  
# Explanation:
# - (?i) -> Case insensitive search:
# - Current: -> Matches "Current:"
# - \\s? -> Matches optional whitespace after the word 'current'
# - (\\d{1,4}(\\.\\d{1,2})?) -> Captures up to 4 digits (e.g., 12, 1234, or 12.34). Only using 4 digits as it is unlikely to see an average greater than xx.xx inches.

# Extract the first match
#combined_text → Contains all the extracted text from the PDF as a single string.
#pattern → The regex pattern we created to find the precipitation value.
#str_match() → Searches for the first occurrence that matches the pattern and extracts the specific value.

match <- str_match(combined_text, pattern)

# Check if a match was found
#the if statement prevents the code from breaking in case there is no match for 'current precipitation' 
if (!is.na(match[2])) {
  N_sierra_current_precip <- match[2]
  cat(paste(N_sierra_current_precip, "inches\n"))
} else {
  cat("No precipitation data found.\n")
}


```

Percent of average rain for this date

```{r}
# Create an empty vector to store the extracted value
N_sierra_average_precip <- c()

# Extract text from the PDF file
#northern_pdf_text <- pdf_text(northern_pdf_file)

# Check if the PDF text is not empty
if (length(northern_pdf_text) > 0) {
  # Concatenate all the text into a single string
  combined_text <- paste(northern_pdf_text, collapse = " ")
  
  # Define the pattern to match the "Percent of Average for this Date:" followed by the percentage value
  pattern <- "(?i)Percent of Average for this Date:\\s*(\\d+%)"
  
  # Extract the match using regmatches and regexpr
  northern_next_four <- regmatches(combined_text, regexpr(pattern, combined_text, perl = TRUE))
  
  if (length(northern_next_four) > 0) {
    # Extract the percentage value using a capturing group
    percent_value <- sub("(?i)Percent of Average for this Date:\\s*(\\d+%)", "\\1", northern_next_four)
    
    # Append the value to the vector
    N_sierra_average_precip <- c(N_sierra_average_precip, percent_value)
    
    # Print only the percentage value
    cat(percent_value, "\n")
  } else {
    cat("Percent of Average for this Date not found.\n")
  }
} else {
  cat("No text extracted from the PDF.\n")
}



```

#### Central Sierra Region Current precipitation {style="color: darkblue"}

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

#Central Sierra precipitation #Pulling and downloading the pdf from online
#URL of pdf file


# URL of the PDF file
central_ppt_pdf_url <- "https://cdec.water.ca.gov/cgi-progs/products/PLOT_FSI.pdf"

# Read PDF text directly from the URL
central_pdf_text <- pdf_text(central_ppt_pdf_url)

# Print the extracted text (for debugging)
print(central_pdf_text)


```

```{r}
# Create an empty vector
Central_current_precip <- c()

# Adjusted pattern for "Current:"
if (length(central_pdf_text) > 0) {
  for (sentence in central_pdf_text) {
    # Pattern to match "Current:" and capture the 4 characters after it
    pattern <- "(?i)Current[:\\s]+(.{4})"
    
    # Extract the matched text
    next_four <- regmatches(sentence, regexpr(pattern, sentence, perl = TRUE))
    
    if (length(next_four) > 0) {
      # Extract only the captured group (the 4 characters) and not "Current:"
      number_only <- sub(pattern, "\\1", next_four)
      
      # Append the value of number_only to the vector
      Central_current_precip <- c(Central_current_precip, number_only)
      
      # Print only the number
      cat(paste(number_only, "inches\n"))
    } else {
      cat("Next four characters after 'Current:' not found.\n")
    }
  }
} else {
  cat("No text extracted from the PDF.\n")
}


```

Central Sierra Region: Percent of average rain for this date

```{r}

#Central Sierra #PUlling out Percent average

#create an empty vector
Central_average_precip <- c()

# Adjusted pattern for "Percent of Average for this Date:"
if (length(central_pdf_text) > 0) {
  for (sentence in central_pdf_text) {
    pattern <- "(?i)Percent of Average for this Date[:\\s]+(.{4})"
    next_four <- regmatches(sentence, regexpr(pattern, sentence, perl = TRUE))
    if (length(next_four) > 0) {
      # Append the value of next_four to the vector
      Central_average_precip <- c(Central_average_precip, next_four)
      cat(paste(next_four, "\n"))
    } else {
      cat("Next four characters after 'Percent of Average for this Date:' not found.\n")
    }
  }
} else {
  cat("No text extracted from the PDF.")
}



```

#################################################################################### 

# Snow Pack

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
#URL of pdf file

# ***** need to see if this pdf updates everyday with the same URL *******
snow_pack_pdf_url <- "https://cdec.water.ca.gov/cgi-progs/products/swccond.pdf"


#pathway to where you want the pdf to be saved
#can eventaully make the name of the pdf with the date it is downloaded
snow_pack_pdf_file<-"data_output/snow_pack.pdf"

# Download the PDF file

# The "wb" mode is typically used when downloading non-text files, such as images, PDFs, or other binary files. It tells R to open the file in binary mode, which is necessary for correctly handling binary data
download.file(snow_pack_pdf_url, snow_pack_pdf_file, mode = "wb")

# Extract text from the PDF
snow_pack_pdf_text <- pdf_text(snow_pack_pdf_file)

# Print the extracted text
print(snow_pack_pdf_text)


```

#### Northern Sierra

```{r}
# Define the pattern to match the line containing "NORTH" and the associated percentages
pattern <- "(?i)NORTH\\s{2,}(\\d+%).*?(\\d+%)"

# Extract the "NORTH" section
north_match <- regmatches(snow_pack_pdf_text, regexec(pattern, snow_pack_pdf_text, perl = TRUE))

# Check if the "NORTH" section is found
if (length(north_match) > 0) {
  # Extract both percentages from the matched text
  north_percentages <- regmatches(north_match[[1]], gregexpr("\\d+%", north_match[[1]]))[[1]]
  
  # Print the extracted percentage
  if (length(north_percentages) >= 2) {
    cat(paste("Northern CA:", north_percentages[2], "of percent avgerage to date snow pack"))
  } else {
    cat("Second percentage in the 'NORTH' section not found.")
  }
} else {
  cat("NORTH section not found.")
}

```

```{r}

# Define the pattern to match the line containing the snow water equivalent information
pattern <- "(?i)Average snow water equivalent \\(Inches\\)\\s+(\\d+\\.\\d+)"

# Extract the line containing snow water equivalent information
snow_water_line <- regmatches(snow_pack_pdf_text, regexpr(pattern, snow_pack_pdf_text, perl = TRUE))

# Check if the line is found
if (length(snow_water_line) > 0) {
  # Extract the numeric value
  numeric_value <- sub("(?i)Average snow water equivalent \\(Inches\\)\\s+(\\d+\\.\\d+)", "\\1", snow_water_line)
  
  # Create the modified line with "inches" after the numeric value
  modified_snow_water_line <- paste("average snow water equivalent", numeric_value, "inches")
  
  cat("Northern Sierra", modified_snow_water_line, "\n")
} else {
  cat("Snow water equivalent information not found.\n")
}


```

#### Central Sierra

```{r}
# Define the pattern to match the line containing "NORTH" and the associated percentages
pattern <- "(?i)CENTRAL SIERRA\\s{2,}(\\d+%).*?(\\d+%)"

# Extract the "CENTRAL SIERRA" section
central_match <- regmatches(snow_pack_pdf_text, regexec(pattern, snow_pack_pdf_text, perl = TRUE))

# Check if the "CENTRAL SIERRA" section is found
if (length(central_match) > 0) {
  # Extract both percentages from the matched text
  central_percentages <- regmatches(central_match[[1]], gregexpr("\\d+%", central_match[[1]]))[[1]]
  
  # Print the extracted percentage
  if (length(central_percentages) >= 2) {
    cat(paste("CENTRAL SIERRA", central_percentages[2], "of percent avg. to date snow pack"))
  } else {
    cat("Second percentage in the 'CENTRAL SIERRA' section not found.")
  }
} else {
  cat("CENTRAL SIERRA section not found.")
}


```

```{r}


# Find the position of "Central Sierra" in the text
central_sierra_position <- regexpr("(?i)Central Sierra", snow_pack_pdf_text)

# If "Central Sierra" is found, find the position of "Average snow water equivalent (Inches)" after it
if (central_sierra_position > 0) {
  # Get the substring starting from the position of "Central Sierra"
  substring_after_central <- substring(snow_pack_pdf_text, central_sierra_position)
  
  # Find the position of "Average snow water equivalent (Inches)" within the substring
  snow_water_position <- regexpr("(?i)Average snow water equivalent \\(Inches\\)", substring_after_central)
  
  # If "Average snow water equivalent (Inches)" is found, extract the numeric value following it
  if (snow_water_position > 0) {
    # Get the substring starting from the position of "Average snow water equivalent (Inches)"
    snow_water_substring <- substring(substring_after_central, snow_water_position)
    
    # Extract the numeric value following "Average snow water equivalent (Inches)"
    snow_water_value <- regmatches(snow_water_substring, regexpr("\\d+\\.\\d+", snow_water_substring))
    
    # If a numeric value is found, print it with the desired text
    if (length(snow_water_value) > 0) {
      cat(paste("Central snow water equivalent", snow_water_value))
    } else {
      cat("Snow water equivalent value not found.")
    }
  } else {
    cat("Average snow water equivalent information not found for Central Sierra.")
  }
} else {
  cat("Central Sierra information not found.")
}


```

#### State-wide

```{r}

# Find the position of "State" in the text
state_position <- regexpr("(?i)State", snow_pack_pdf_text)

# If "State" is found, find the position of the numeric value after it
if (state_position > 0) {
  # Get the substring starting from the position of "State"
  substring_after_state <- substring(snow_pack_pdf_text, state_position)
  
  # Extract the numeric value following "State"
  snow_water_value <- regmatches(substring_after_state, regexpr("\\d+\\.\\d+", substring_after_state))
  
  # If a numeric value is found, print it with the desired text
  if (length(snow_water_value) > 0) {
    cat(paste("State wide average snow pack", snow_water_value))
  } else {
    cat("Snow water equivalent value not found.")
  }
} else {
  cat("State information not found.")
}


```

```{r}

# Find the position of "Statewide Average:" in the text
statewide_position <- regexpr("(?i)Statewide Average:", snow_pack_pdf_text)

# If "Statewide Average:" is found, find the position of the numeric value after it
if (statewide_position > 0) {
  # Get the substring starting from the position of "Statewide Average:"
  substring_after_statewide <- substring(snow_pack_pdf_text, statewide_position)
  
  # Extract both percentages from the matched text
  statewide_percentages <- regmatches(substring_after_statewide, gregexpr("\\d+%", substring_after_statewide))[[1]]
  
  # Print the second percentage with the desired text
  if (length(statewide_percentages) >= 2) {
    cat("Statewide average percent snow pack for this date", statewide_percentages[2], "\n")
  } else {
    cat("Second percentage after 'Statewide Average:' not found.")
  }
} else {
  cat("Statewide Average information not found.\n")
}



```

# Reservoir Storage

```{r}
file.exists("geo_files/legal_delta.geojson")
# Load the Legal Delta boundary (replace with your actual file path)
legal_delta_boundary <- geojson_read("geo_files/legal_delta.geojson", what = "sp")

# Create the map with Legal Delta boundary and specific points
leaflet() %>%
  addTiles() %>%
  setView(lng = -121.5, lat = 38.05, zoom = 7) %>%  # Center the map

  # Add the Legal Delta boundary polygon
  addPolygons(data = legal_delta_boundary, color = "blue", weight = 2, fillOpacity = 0.5, popup = "Legal Delta Boundary") %>%
  
  # Add specific points of interest as markers
  addMarkers(lng = -121.493000, lat = 39.540000, popup = "Oroville Reservoir") %>%  # Oroville reservoir (ORO)
  addMarkers(lng = 	-119.302000, lat = 	37.145000, popup = "Shasta Reservoir") %>%    # Shasta reservoir (SHA)
  addMarkers(lng = -121.133000, lat = 37.033000, popup = "San Luis Reservoir")   # San Luis reservoir (SNL)



```

##### shaded area on map is the legal delta

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

# URL of the website to scrape
reservoir_url <- "https://cdec.water.ca.gov/reportapp/javareports?name=RES"

# Read the web page
reservoir_web_page <- read_html(reservoir_url)

# Extract the table node
reservoir_table_node <- html_node(reservoir_web_page, "table")

# Extract the table content into a data frame
reservoir_table_data <- html_table(reservoir_table_node)

# Print the entire table data
print(reservoir_table_data)

```

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
# Extract a specific numbers [row, column]
# Shasta (at Sacramento River)

# extract percent of average storage
shasta_percent_average_storage_value <- as.numeric(reservoir_table_data[10, 9])

# extract percent of capacity
shasta_percent_capacity_value <- as.numeric(reservoir_table_data[10, 7])


# Print the formatted output with a Markdown header
cat(sprintf(" Shasta Reservoir: %.0f%% of average storage, %.0f%% of total capacity.\n", 
    shasta_percent_average_storage_value, shasta_percent_capacity_value))


```

### Shasta Reservoir: `r shasta_percent_average_storage_value`% of average storage, `r shasta_percent_capacity_value`% of total capacity.

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
# Extract a specific numbers [row, column]
# Oroville (at Feather River)

# extract percent of average storage
oroville_percent_average_storage_value <- as.numeric(reservoir_table_data[13, 9])

# extract percent of capacity
oroville_percent_capacity_value <- as.numeric(reservoir_table_data[13, 7])

# Print the formatted output (but this will be hidden)
cat(sprintf(" Oroville Reservoir: %.0f%% of average storage, %.0f%% of total capacity.\n", 
    oroville_percent_average_storage_value, oroville_percent_capacity_value))

```

### Oroville Reservoir: `r oroville_percent_average_storage_value`% of average storage, `r oroville_percent_capacity_value`% of total capacity.

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

# Extract a specific numbers [row, column]
# San Luis (at San Luis Creek)

# extract percent of average storage
SanLuis_percent_average_storage_value <- as.numeric(reservoir_table_data[56, 9])

# extract percent of capacity
SanLuis_percent_capacity_value <- as.numeric(reservoir_table_data[56, 7])

# Print the formatted output (but this will be hidden)
cat(sprintf(" San Luis Reservoir: %.0f%% of average storage, %.0f%% of total capacity.\n", 
    SanLuis_percent_average_storage_value, SanLuis_percent_capacity_value))

```

### San Luis Reservoir: `r SanLuis_percent_average_storage_value`% of average storage, `r SanLuis_percent_capacity_value`% of total capacity.

# Water Flow and Quality

#### Location of flow sensors. Shaded region is legal delta

```{r}
#map of water flow sensors

# Create the map with Legal Delta boundary and specific points
leaflet() %>%
  addTiles() %>%
  setView(lng = -121.5, lat = 38.05, zoom = 8) %>%
  
  addPolygons(data = legal_delta_boundary, color = "blue", weight = 2, fillOpacity = 0.5, popup = "Legal Delta Boundary") %>%
  
  addMarkers(lng = -121.27, lat = 37.68, popup = "San Joaquin River near Vernalis") %>% # Add specific gps points
  addMarkers(lng = -121.7, lat = 38.45, popup = "Sacramento at Freeport") %>%          # Add specific gps points
  addMarkers(lng = -121.3, lat = 37.9, popup = "Central Valley Project and State Water Project pumping station")   # Center the map



             


```

#### Sacramento River at Freeport

```{r}


#### Sacramento River at Freeport

# File path to your Excel file
sacramento_file_path <- "historical_data/historical_data_full.xlsx"

# Determine the current month and year
sacramento_current_month <- format(Sys.Date(), "%B")
sacramento_current_year <- format(Sys.Date(), "%Y")

# Load the entire workbook without overwriting other sheets
sacramento_workbook <- loadWorkbook(sacramento_file_path)

# Read the specific sheet you want to update
sacramento_historical_data <- readWorkbook(sacramento_workbook, sheet = "Flow_Sac")

# Ensure Year is character
sacramento_historical_data$Year <- as.character(sacramento_historical_data$Year) 

# Define the start and end dates for the current month
sacramento_start_date <- floor_date(Sys.Date(), "month")
sacramento_end_date <- Sys.Date()

# Query CDEC data for Sacramento at Freeport (station "FPT") for the current month
sacramento_flow_daily <- cdec_query(
  station = "FPT", 
  sensor_num = 20, 
  dur_code = "H", 
  start_date = as.character(sacramento_start_date), 
  end_date = as.character(sacramento_end_date)
)

# Calculate the monthly average for Sacramento flow
sacramento_export_monthly_average <- mean(sacramento_flow_daily$parameter_value, na.rm = TRUE)

# Calculate the historical mean for the current month, if available in your data
sacramento_current_month_historical_data <- sacramento_historical_data %>%
  filter(Month == sacramento_current_month) %>%
  select(Average_Flow_Daily)

sacramento_current_month_historical_mean <- mean(sacramento_current_month_historical_data$Average_Flow_Daily, na.rm = TRUE)

# Calculate the percentage of the current month’s average relative to the historical average
sacramento_percentage_of_cfs_average <- (sacramento_export_monthly_average / sacramento_current_month_historical_mean) * 100




# Define HTML color codes for Quarto rendering (for percentage only)
sac_color_percentage <- ifelse(
  sacramento_percentage_of_cfs_average >= 90, "green",
  ifelse(sacramento_percentage_of_cfs_average >= 75, "blue",
  ifelse(sacramento_percentage_of_cfs_average >= 50, "orange",
  "red"))) 


sac_flow_formatted <- sprintf(
  "Sacramento export monthly average is %.2f%% of the historical average for %s.<br>
   Average CFS for %s %s is <strong><span style='color:%s'>%.2f</span></strong>.<br>", 
  sacramento_percentage_of_cfs_average, sacramento_current_month, 
  sacramento_current_month, sacramento_current_year, 
  sac_color_percentage, sacramento_export_monthly_average
)

# Render as raw HTML so Quarto interprets it correctly
HTML(sac_flow_formatted)

# Check if there is an entry for the current month and year
existing_entry <- sacramento_historical_data %>%
  filter(Month == sacramento_current_month, Year == sacramento_current_year)

suppressMessages({
  if (nrow(existing_entry) > 0) {  
    # Update the existing row with the new flow average for the current month
    row_to_update <- which(
      sacramento_historical_data$Month == sacramento_current_month & 
      sacramento_historical_data$Year == sacramento_current_year
    )
    
    sacramento_historical_data[row_to_update, "Average_Flow_Daily"] <- sacramento_export_monthly_average
  } else { 
    # Create a new row with the current month, year, and flow average if it's a new month
    new_row <- data.frame(
      Month = sacramento_current_month, 
      Year = sacramento_current_year, 
      Average_Flow_Daily = sacramento_export_monthly_average
    )
    
    # Append the new row to the historical data
    sacramento_historical_data <- bind_rows(sacramento_historical_data, new_row)
  }

  # Write only the modified sheet back to the workbook
  writeData(sacramento_workbook, sheet = "Flow_Sac", x = sacramento_historical_data)

  # Save the workbook, preserving all other sheets
  saveWorkbook(sacramento_workbook, sacramento_file_path, overwrite = TRUE)
})


```


```{r}

#Sacramento River at Hood (station code SRH) #When run, will automatically update for current month

current_month <- format(Sys.Date(), "%B")

# Get the first day of the current month
start_date <- floor_date(Sys.Date(), "month")

# Get the current date
end_date <- Sys.Date()

# Query CDEC data from the first of the month to the current date
sac_cond_daily_new <- cdec_query(station = "SRH", sensor_num = "100", 
                             dur_code = "D", start_date = as.character(start_date), 
                             end_date = as.character(end_date))
Sac_cond_monthly_average<-mean(sac_cond_daily_new$parameter_value, na.rm = T)


#### ppm conversion
sac_ppm<-(Sac_cond_monthly_average*0.55)



#output
sac_salinity_formatted<-sprintf("Average monthly salinity for %s is %.2f µS/cm | %.2f ppm",
                current_month, Sac_cond_monthly_average, sac_ppm)

HTML(sac_salinity_formatted)
```


#### San Joaquin River at Vernalis

```{r}

# File path to your Excel file (the one you just uploaded)
san_joaquin_file_path <- "historical_data/historical_data_full.xlsx"

# Determine the current month and year
san_joaquin_current_month <- format(Sys.Date(), "%B")
san_joaquin_current_year <- format(Sys.Date(), "%Y")

# Load the entire workbook without overwriting other sheets
san_joaquin_workbook <- loadWorkbook(san_joaquin_file_path)

# Read the specific sheet you want to update
san_joaquin_historical_data <- readWorkbook(san_joaquin_workbook, sheet = "Flow_SanJoaq")
san_joaquin_historical_data$Year <- as.character(san_joaquin_historical_data$Year) # Ensure Year is character

# Define the start and end dates for the current month
san_joaquin_start_date <- floor_date(Sys.Date(), "month")
san_joaquin_end_date <- Sys.Date()

# Query CDEC data for San Joaquin at Vernalis (station "VNS") for the current month
# Assuming `cdec_query` returns a data frame with a "parameter_value" column
san_joaquin_flow_daily <- cdec_query(station = "VNS", sensor_num = 20, 
                                     dur_code = "H", start_date = as.character(san_joaquin_start_date), 
                                     end_date = as.character(san_joaquin_end_date))

# Calculate the monthly average for San Joaquin flow
san_joaquin_export_monthly_average <- mean(san_joaquin_flow_daily$parameter_value, na.rm = TRUE)

# Calculate the historical mean for the current month, if available in your data
san_joaquin_current_month_historical_data <- san_joaquin_historical_data %>%
  filter(Month == san_joaquin_current_month) %>%
  select(Average_Flow_Daily)

san_joaquin_current_month_historical_mean <- mean(san_joaquin_current_month_historical_data$Average_Flow_Daily, na.rm = TRUE)

# Calculate the percentage of the current month’s average relative to the historical average
san_joaquin_percentage_of_cfs_average <- (san_joaquin_export_monthly_average / san_joaquin_current_month_historical_mean) * 100




# Define HTML color codes for Quarto rendering (for percentage only)
sanjoaq_color_percentage <- ifelse(
  san_joaquin_percentage_of_cfs_average >= 90, "green",
  ifelse(san_joaquin_percentage_of_cfs_average >= 75, "blue",
  ifelse(san_joaquin_percentage_of_cfs_average >= 50, "orange",
  "red")))


sanjoaq_flow_formatted <- sprintf(
  "San Joaquin export monthly average is %.2f%% of the historical average for %s.<br>
   Average CFS for %s %s is <strong><span style='color:%s'>%.2f</span></strong>.<br>", 
 san_joaquin_percentage_of_cfs_average, san_joaquin_current_month, 
 san_joaquin_current_month, san_joaquin_current_year, 
  sanjoaq_color_percentage, san_joaquin_export_monthly_average
)

# Render as raw HTML so Quarto interprets it correctly
HTML(sanjoaq_flow_formatted)





# Check if there is an entry for the current month and year
existing_entry <- san_joaquin_historical_data %>%
  filter(Month == san_joaquin_current_month, Year == san_joaquin_current_year)

suppressMessages({
  if (nrow(existing_entry) > 0) {
    # Update the existing row with the new flow average for the current month
    row_to_update <- which(san_joaquin_historical_data$Month == san_joaquin_current_month & 
                           san_joaquin_historical_data$Year == san_joaquin_current_year)
    san_joaquin_historical_data[row_to_update, "Average_Flow_Daily"] <- san_joaquin_export_monthly_average
  } else {
    # Create a new row with the current month, year, and flow average if it's a new month
    new_row <- data.frame(
      Month = san_joaquin_current_month,
      Year = san_joaquin_current_year,
      Average_Flow_Daily = san_joaquin_export_monthly_average
    )
    
    # Append the new row to the historical data
    san_joaquin_historical_data <- bind_rows(san_joaquin_historical_data, new_row)
  }
  
  # Write only the modified sheet back to the workbook
  writeData(san_joaquin_workbook, sheet = "Flow_SanJoaq", x = san_joaquin_historical_data)
  
  # Save the workbook, preserving all other sheets
  saveWorkbook(san_joaquin_workbook, san_joaquin_file_path, overwrite = TRUE)
})



```

```{r}

#San Joaquin River at Vernalis (station code - VER) #When run, will automatically update for current month

# Get the first day of the current month
start_date <- floor_date(Sys.Date(), "month")

# Get the current date
end_date <- Sys.Date()

# Query CDEC data from the first of the month to the current date
SanJoaquin_cond_daily_new <- cdec_query(station = "VER", sensor_num = "100", 
                             dur_code = "D", start_date = as.character(start_date), 
                             end_date = as.character(end_date))


SanJoaquin_cond_monthly_average<-mean(SanJoaquin_cond_daily_new$parameter_value, na.rm = T)


#### ppm conversion
SanJoa_ppm<-(SanJoaquin_cond_monthly_average*0.55)


#output



#output
SanJoa_salinity_formatted<-sprintf("Average monthly salinity for %s is %.2f µS/cm | %.2f ppm",
                current_month, SanJoaquin_cond_monthly_average, SanJoa_ppm)

HTML(SanJoa_salinity_formatted)

```

#### Total Delta Exports: (Central Valley Project + State Water Project):

This is a monthly average of exports through: Clifton Court (CLT), Tracy (TRA), Contra Costa (CCC), and Byron Bethany (BBID)

```{r}

# URL of the PDF file
delta_exports_pdf_url <- "https://www.usbr.gov/mp/cvo/vungvari/doutdly.pdf"

# Path to save the downloaded PDF file
delta_exports_pdf_file <- "historical_data/delta_exports.pdf"

# Download the PDF file
download.file(delta_exports_pdf_url, delta_exports_pdf_file, mode = "wb")

# Extract text from the PDF file
delta_exports_pdf_text <- pdf_text(delta_exports_pdf_file)

# Extract the last page's text
delta_export_last_page_text <- delta_exports_pdf_text[length(delta_exports_pdf_text)]

# Split the text into lines
delta_export_lines <- str_split(delta_export_last_page_text, "\\n")[[1]]

# Find the line containing the "Average" row
delta_export_average_line <- delta_export_lines[str_detect(delta_export_lines, "Average")]

# Define delta_export_numbers before use
delta_export_numbers <- str_split(delta_export_average_line, "\\s+")[[1]]

# Print extracted elements for debugging
#print(delta_export_numbers)

# Extract numeric values
numeric_values <- delta_export_numbers[str_detect(delta_export_numbers, "^\\d{1,5}[,\\.]?\\d*$")]

# Print numeric values for debugging
#print(numeric_values)

# Extract the 17th numeric value

#sometimes the number (length(numeric_values) >= 15) needs to be changed 

if (length(numeric_values) >= 15) {
    average_delta_export_value <- as.numeric(gsub(",", "", numeric_values[15]))
} else {
    warning("Could not find the 14th numeric value. Check the PDF format.")
    average_delta_export_value <- NA
}

# Print the final extracted value
#print(average_delta_export_value)

# Excel file path and current date
delta_export_file_path <- "historical_data/historical_data_full.xlsx"
current_month <- format(Sys.Date(), "%B")
current_year <- format(Sys.Date(), "%Y")

# Load workbook and sheet
delta_exports_workbook <- loadWorkbook(delta_export_file_path)
delta_exports_data <- readWorkbook(delta_exports_workbook, sheet = "Total_Delta_Exports")

delta_exports_data$Year <- as.character(delta_exports_data$Year)

# Check for existing entry
existing_entry <- delta_exports_data %>%
  filter(Month == current_month, Year == current_year)

if (nrow(existing_entry) > 0) {
  row_to_update <- which(delta_exports_data$Month == current_month &
                         delta_exports_data$Year == current_year)
  delta_exports_data[row_to_update, "Exports"] <- average_delta_export_value
} else {
  new_row <- data.frame(
    Month = current_month,
    Year = current_year,
    Exports = average_delta_export_value
  )
  delta_exports_data <- bind_rows(delta_exports_data, new_row)
}

# Write updated sheet back to the workbook
writeData(delta_exports_workbook, sheet = "Total_Delta_Exports", x = delta_exports_data)

# Save the workbook
saveWorkbook(delta_exports_workbook, delta_export_file_path, overwrite = TRUE)

####################

# Determine the current month and year
delta_export_current_month <- format(Sys.Date(), "%B")
delta_export_current_year <- format(Sys.Date(), "%Y")

# Load the entire workbook without overwriting other sheets
delta_export_workbook <- loadWorkbook(delta_export_file_path)

# Read the specific sheet you want to update
delta_export_historical_data <- readWorkbook(delta_export_workbook, sheet = "Total_Delta_Exports")
delta_export_historical_data$Year <- as.character(delta_export_historical_data$Year) # Ensure Year is character

# Calculate the historical mean for the current month, if available in your data
delta_export_current_month_historical_data <- delta_export_historical_data %>%
  filter(Month == delta_export_current_month) %>%
  select(Exports)


delta_export_current_month_historical_mean <- mean(delta_export_current_month_historical_data$Exports, na.rm = TRUE)
#delta_export_current_month_historical_mean


# Calculate the percentage of the current month’s average relative to the historical average
delta_export_percentage_of_cfs_average <- (average_delta_export_value / delta_export_current_month_historical_mean) * 100
#delta_export_percentage_of_cfs_average


#.2f specifies 2 digits after the decimal place, and f stands for floating point number - meaning the decimal point will be placed whereever it is needed




# Define HTML color codes for Quarto rendering (for percentage only)
deltaexport_color_percentage <- ifelse(
  delta_export_percentage_of_cfs_average >= 90, "green",
  ifelse(delta_export_percentage_of_cfs_average >= 75, "blue",
  ifelse(delta_export_percentage_of_cfs_average >= 50, "orange",
  "red")))


delta_export_formatted <- sprintf(
  "Total Delta export (CVP + SWP) monthly average is %.2f%% of the historical average for %s.<br>
   Average CFS for %s %s is <strong><span style='color:%s'>%.2f</span></strong>.<br>", 
 delta_export_percentage_of_cfs_average, delta_export_current_month, 
 delta_export_current_month, delta_export_current_year, 
  deltaexport_color_percentage, average_delta_export_value
)

# Render as raw HTML so Quarto interprets it correctly
HTML(delta_export_formatted)



```

```{r}
#Salinity at Banks pumping plant (station code - HBP) #When run, will automatically update for current month

# Get the first day of the current month
start_date <- floor_date(Sys.Date(), "month")

# Get the current date
end_date <- Sys.Date()


#### this sensor is not online at the moment
# Query CDEC data from the first of the month to the current date
Banks_cond_daily_new <- cdec_query(station = "HBP", sensor_num = "100", 
                             dur_code = "D", start_date = as.character(start_date), 
                             end_date = as.character(end_date))



#need to take the average of all parameter_value for the date 
Banks_cond_monthly_average<-mean(Banks_cond_daily_new$parameter_value, na.rm = T)



#convert um/cm to ppm
#idea to have some historical context with the salinity?
#### ppm conversion
Banks_ppm<-(Banks_cond_monthly_average*0.55)



#output
banks_salinity_formatted<-sprintf("Average monthly salinity for %s is %.2f µS/cm | %.2f ppm",
                current_month, Banks_cond_monthly_average, Banks_ppm)

HTML(banks_salinity_formatted)


```

```{r}
# I want to put the color coding for the sets of numbers here

# Define the ranges with actual numeric values
ranges <- c("<span style='color:red'>50%</span>", 
            "<span style='color:orange'>50-74%</span>", 
            "<span style='color:blue'>75-90%</span>", 
            "<span style='color:green'>>90%</span>")

# Create the inline HTML string with color-coded numbers
html_string <- paste(ranges, collapse = "  ")

# Render as raw HTML in Quarto
cat(html_string)


```



# TEMPERATURE

#need to add hisotrical average data in --\> will add to csv file eventually

```{r}

#Water Temp - Sacramento river headwaters at Clear Creek

# Get the first day of the current month
start_date <- floor_date(Sys.Date(), "month")

# Get the current date
end_date <- Sys.Date()

# Query CDEC data from the first of the month to the current date
ClearCreek_waterTemp_new <- cdec_query(station = "IGO", sensor_num = "25", 
                             dur_code = "H", start_date = as.character(start_date), 
                             end_date = as.character(end_date))



#need to take the average of all parameter_value for the date 
ClearCreek_waterTemp_monthly_average<-mean(ClearCreek_waterTemp_new$parameter_value, na.rm = T)




#cdec_datasets("IGO")




#### write river temp into workbook

# Excel file path and current date
ClearCreek_file_path <- "historical_data/historical_data_full.xlsx"
current_month <- format(Sys.Date(), "%B")
current_year <- format(Sys.Date(), "%Y")

# Load workbook and sheet
ClearCreek_workbook <- loadWorkbook(ClearCreek_file_path)
ClearCreek_data <- readWorkbook(ClearCreek_workbook, sheet = "Temp_IGO")

ClearCreek_data$Year <- as.character(ClearCreek_data$Year)

# Check for existing entry
existing_entry <- ClearCreek_data %>%
  filter(Month == current_month, Year == current_year)

if (nrow(existing_entry) > 0) {
  row_to_update <- which(ClearCreek_data$Month == current_month &
                         ClearCreek_data$Year == current_year)
  ClearCreek_data[row_to_update, "Average_Temp_Monthly"] <- ClearCreek_waterTemp_monthly_average
} else {
  new_row <- data.frame(
    Month = current_month,
    Year = current_year,
    Average_Temp_Monthly = ClearCreek_waterTemp_monthly_average
  )
  ClearCreek_data <- bind_rows(ClearCreek_data, new_row)
}

# Write updated sheet back to the workbook
writeData(ClearCreek_workbook, sheet = "Temp_IGO", x = ClearCreek_data)

# Save the workbook
saveWorkbook(ClearCreek_workbook, ClearCreek_file_path, overwrite = TRUE)

##### make it print the monthly average for that month for a 20 year average (i think its actually 15 for these)
######################

# Determine the current month and year
ClearCreek_current_month <- format(Sys.Date(), "%B")
ClearCreek_current_year <- format(Sys.Date(), "%Y")

# Load the entire workbook without overwriting other sheets
ClearCreek_workbook <- loadWorkbook(ClearCreek_file_path)

# Read the specific sheet you want to update
ClearCreek_historical_data <- readWorkbook(ClearCreek_workbook, sheet = "Temp_IGO")
ClearCreek_historical_data$Year <- as.character(ClearCreek_historical_data$Year) # Ensure Year is character

# Calculate the historical mean for the current month, if available in your data
ClearCreek_current_month_historical_data <- ClearCreek_historical_data %>%
  filter(Month == ClearCreek_current_month) %>%
  select(Average_Temp_Monthly)


ClearCreek_current_month_historical_mean <- mean(ClearCreek_current_month_historical_data$Average_Temp_Monthly, na.rm = TRUE)
#delta_export_current_month_historical_mean





ClearCreek_temp_formatted<-sprintf("Average monthly Temperature at Clear Creek for %s is %.2f °F and the historical monthly average temperature is %.2f °F",
                ClearCreek_current_month, ClearCreek_waterTemp_monthly_average,ClearCreek_current_month_historical_mean)

HTML(ClearCreek_temp_formatted)


```

```{r}
#Water Temp - Sacramento/San Joaquin Rivers confluence at Collinsville

# Get the first day of the current month
start_date <- floor_date(Sys.Date(), "month")

# Get the current date
end_date <- Sys.Date()

# Query CDEC data from the first of the month to the current date
SacSanJoaquin_waterTemp_new <- cdec_query(station = "CSE", sensor_num = "25", 
                             dur_code = "H", start_date = as.character(start_date), 
                             end_date = as.character(end_date))



#need to take the average of all parameter_value for the date 
SacSanJoaquin_waterTemp_monthly_average<-mean(SacSanJoaquin_waterTemp_new$parameter_value, na.rm = T)



#### write river temp into workbook

# Excel file path and current date
SacSanJoaquin_file_path <- "historical_data/historical_data_full.xlsx"
current_month <- format(Sys.Date(), "%B")
current_year <- format(Sys.Date(), "%Y")

# Load workbook and sheet
SacSanJoaquin_workbook <- loadWorkbook(SacSanJoaquin_file_path)
SacSanJoaquin_data <- readWorkbook(SacSanJoaquin_workbook, sheet = "Temp_CSE")

SacSanJoaquin_data$Year <- as.character(SacSanJoaquin_data$Year)

# Check for existing entry
existing_entry <- SacSanJoaquin_data %>%
  filter(Month == current_month, Year == current_year)

if (nrow(existing_entry) > 0) {
  row_to_update <- which(SacSanJoaquin_data$Month == current_month &
                         SacSanJoaquin_data$Year == current_year)
  SacSanJoaquin_data[row_to_update, "Average_Temp_Monthly"] <- SacSanJoaquin_waterTemp_monthly_average
} else {
  new_row <- data.frame(
    Month = current_month,
    Year = current_year,
    Average_Temp_Monthly = SacSanJoaquin_waterTemp_monthly_average
  )
  SacSanJoaquin_data <- bind_rows(SacSanJoaquin_data, new_row)
}

# Write updated sheet back to the workbook
writeData(SacSanJoaquin_workbook, sheet = "Temp_CSE", x = SacSanJoaquin_data)

# Save the workbook
saveWorkbook(SacSanJoaquin_workbook, SacSanJoaquin_file_path, overwrite = TRUE)

##### make it print the monthly average for that month for a 20 year average (i think its actually 15 for these)
######################

# Determine the current month and year
SacSanJoaquin_current_month <- format(Sys.Date(), "%B")
SacSanJoaquin_current_year <- format(Sys.Date(), "%Y")

# Load the entire workbook without overwriting other sheets
SacSanJoaquin_workbook <- loadWorkbook(SacSanJoaquin_file_path)

# Read the specific sheet you want to update
SacSanJoaquin_historical_data <- readWorkbook(SacSanJoaquin_workbook, sheet = "Temp_CSE")
SacSanJoaquin_historical_data$Year <- as.character(SacSanJoaquin_historical_data$Year) # Ensure Year is character

# Calculate the historical mean for the current month, if available in your data
SacSanJoaquin_current_month_historical_data <- SacSanJoaquin_historical_data %>%
  filter(Month == SacSanJoaquin_current_month) %>%
  select(Average_Temp_Hourly)


SacSanJoaquin_current_month_historical_mean <- mean(SacSanJoaquin_current_month_historical_data$Average_Temp_Hourly, na.rm = TRUE)
#delta_export_current_month_historical_mean




SacSanJoaquin_temp_formatted<-sprintf("Average monthly Temperature at Collinsville for %s is %.2f °F and the historical monthly average temperature is %.2f °F",
                SacSanJoaquin_current_month, SacSanJoaquin_waterTemp_monthly_average, SacSanJoaquin_current_month_historical_mean)

HTML(SacSanJoaquin_temp_formatted)


```

```{r}

#Water Temp - South Delta (near HBP) at Clifton Court Forebay


#cdec_datasets("CLC")

# Get the first day of the current month
start_date <- floor_date(Sys.Date(), "month")

# Get the current date
end_date <- Sys.Date()

# Query CDEC data from the first of the month to the current date
SouthDelta_waterTemp_new <- cdec_query(station = "CLC", sensor_num = "146", 
                             dur_code = "H", start_date = as.character(start_date), 
                             end_date = as.character(end_date))



#because CLC temp gage only have readings in Celcius, need to convert to F

SouthDelta_waterTemp_new$parameter_F<-(SouthDelta_waterTemp_new$parameter_value*( 9/5) + 32)


#need to take the average of all parameter_value for the date 
SouthDelta_waterTemp_monthly_average_F<-mean(SouthDelta_waterTemp_new$parameter_F, na.rm = T)






#### write river temp into workbook

# Excel file path and current date
SouthDelta_file_path <- "historical_data/historical_data_full.xlsx"
current_month <- format(Sys.Date(), "%B")
current_year <- format(Sys.Date(), "%Y")

# Load workbook and sheet
SouthDelta_workbook <- loadWorkbook(SouthDelta_file_path)
SouthDelta_data <- readWorkbook(SouthDelta_workbook, sheet = "Temp_CLC")

SouthDelta_data$Year <- as.character(SouthDelta_data$Year)

# Check for existing entry
existing_entry <- SouthDelta_data %>%
  filter(Month == current_month, Year == current_year)

if (nrow(existing_entry) > 0) {
  row_to_update <- which(SouthDelta_data$Month == current_month &
                         SouthDelta_data$Year == current_year)
  SouthDelta_data[row_to_update, "Average_Temp_Monthly"] <- SouthDelta_waterTemp_monthly_average_F
} else {
  new_row <- data.frame(
    Month = current_month,
    Year = current_year,
    Average_Temp_Monthly = SouthDelta_waterTemp_monthly_average_F
  )
  SouthDelta_data <- bind_rows(SouthDelta_data, new_row)
}

# Write updated sheet back to the workbook
writeData(SouthDelta_workbook, sheet = "Temp_CLC", x = SouthDelta_data)

# Save the workbook
saveWorkbook(SouthDelta_workbook, SouthDelta_file_path, overwrite = TRUE)

##### make it print the monthly average for that month for a 20 year average (i think its actually 15 for these)
######################

# Determine the current month and year
SouthDelta_current_month <- format(Sys.Date(), "%B")
SouthDelta_current_year <- format(Sys.Date(), "%Y")

# Load the entire workbook without overwriting other sheets
SouthDelta_workbook <- loadWorkbook(SouthDelta_file_path)

# Read the specific sheet you want to update
SouthDelta_historical_data <- readWorkbook(SouthDelta_workbook, sheet = "Temp_CLC")
SouthDelta_historical_data$Year <- as.character(SouthDelta_historical_data$Year) # Ensure Year is character

# Calculate the historical mean for the current month, if available in your data
SouthDelta_current_month_historical_data <- SouthDelta_historical_data %>%
  filter(Month == SouthDelta_current_month) %>%
  select(Average_Temp_Hourly)


SouthDelta_current_month_historical_mean <- mean(SouthDelta_current_month_historical_data$Average_Temp_Hourly, na.rm = TRUE)
#delta_export_current_month_historical_mean




SouthDelta_temp_formatted<-sprintf("Average monthly Temperature at Clifton Court Forebay for %s is %.2f °F and the historical monthly average temperature is %.2f °F",
                SouthDelta_current_month, SouthDelta_waterTemp_monthly_average_F, SouthDelta_current_month_historical_mean)

HTML(SouthDelta_temp_formatted)


```
